### 2022-08-05 chapter2/chapter.tex
ch 7/8 comments

In this chapter we introduce the discontinuous Galerkin (DG) method applied to fluid queues to approximate the operator-analytic expressions for fluid-fluid queues in \cite{bo2014}. 
\begin{center}
	\begin{minipage}{0.8\textwidth}
		\textit{Apart from Sections~\ref{sec: other applications}-\ref{sec: slope limiting}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis. I am a co-author of the paper \cite{blnos2022}. %The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara. I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme. I contributed Sections~4.4 and 5.1. I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6. Appendix~A is also my original work. I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
		}
	\end{minipage}
	\end{center}
% The DG method for B 
    % direct copy paste from the paper
% Problems with DG, non-neg/oscillations
    % solutions: filtering and limiting
% Approximating R: projection, interpolation, cell averages (eluding to QBD-RAP and unif method)
% Approximating Psi (and the limiting distribution, but this isnt really necessary)
    % constructing D
    % solving the matrix-Riccati equation, iteration/algorithm.

@@

% \section{Discontinuous Galerkin Approximation of the Generator of a Fluid Queue}
	% \label{sec:DG}
Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs). For a most readable and practical text on DG methods we refer the reader to \cite{nodalDGBook}. Here we describe how one can apply the DG method to approximate fluid queues and fluid-fluid queues. In finite-element methods, such as the DG method, the domain of approximation is partitioned into intervals, referred to individually as \textit{cells} and collectively as a \textit{mesh}. On each cell we construct a finite-dimensional smooth Sobolev space using piecewise-polynomial basis functions, and then project the partial differential equations onto this space. This projection leads to a new system of equations which represent a \textit{weak form} of the original system of PDEs. Next, we approximate the \textit{flux operator} which moves probability from one cell to another, in a manner similar to the underlying principle of a finite-volume approximation. This creates a system of ordinary differential equations which is known as a \emph{semi-discrete system} (discrete in the spatial variable but continuous in time). We can solve the semi-discrete system to approximate transient distributions of the fluid queue or, as we do in Section~\ref{sec:DGSFFM}, we can extract and manipulate an approximation to the generator of the fluid queue to approximate performance measures of fluid-fluid queues. The DG method conserves probability. Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the limiting distribution of a stochastic fluid-fluid queue.

@@

For each \(k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) to form a basis for the space \(W_k\), in which we formulate the approximation. Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be the space of polynomials of degree \(p_k-1\). It is convenient in this work to take \(\{\phi^r_k\}_{r=1}^{p_k}\) as a basis of Lagrange interpolating polynomials defined by the Gauss-Lobatto quadrature points, since our approximations inherit nice properties from this \citep{nodalDGBook}. However, the constructions presented here are general, and any basis can be used. For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials. On each cell \(\mathcal D_{k,i}\) we approximate 
\[f_i(x,t)\approx u_{k,i}(x,t)=\sum\limits_{r=1}^{p_k}a_{k,i}^r(t)\phi^r_k(x),\] 
where \(a_{k,i}^r(t)\) are yet-to-be-determined time-dependent coefficients. We refer to \(u_{k,i}\) as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain. The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\mathbb Z\).

@@

The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation~(\ref{eqn:pde_density_matrix}). In general, to construct the weak-form we need a set of test functions, say \(W\). Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to get 
\begin{align}
\begin{multlined}[t]
	\int_{x\in\mathbb R}\cfrac{\partial}{\partial t} f_j(x,t)\psi(x)\wrt x = \int_{x\in\mathbb R} \sum_{i\in\calS}f_i(x,t)T_{ij}\psi(x)\wrt x 
	%
	+  \int_{x\in\mathbb R} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\psi(x)\wrt x \\{}- [ f_j(x,t)c_j\psi(x)]_{x=-\infty}^{x=\infty}, \end{multlined}\label{eqn:weak form}
\end{align}
for \(j\in\calS\). It is common to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero. Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE. For a sufficiently rich set of test functions \(W\) the weak and strong forms of the PDE are equivalent. Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE. For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.

For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\), the same as the set of basis functions of our solution space. Proceeding as described above, the weak formulation is 
\begin{align*}
	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS}  f_i(x,t)T_{ij}\phi^r_k(x)\wrt x  
	%
	+  \int_{x=y_k}^{y_{k+1}} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
\end{align*}
since \(\phi^r_k\) is compactly supported on \(\calD_{k,j}\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\). Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\). Also, note that \(\cfrac{\wrt}{\wrt x}\phi^r_k(x)\in W\). Using this, we can write 
\begin{align*}
	\begin{multlined}[t]\int_{x=y_{k}}^{y_{k+1}}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
	= \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
	%
	\\ {}
	+  \int_{x=y_k}^{y_{k+1}} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
\end{align*}
	which is equivalent to
\begin{align}
	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi^r_k(x)\wrt x  
	%
	\\{} + \int_{x=y_k}^{y_{k+1}} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
\end{align}
Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime, \({}\tr{}\), denotes the transpose, which we now substitute into (\ref{eqn:hash})
%\begin{align*}
%	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi^k(x)\tr{}\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\phi^r_k(x)\wrt x  
%	%
%	\\ {}+  \int_{x\in\calD_k}\boldsymbol a_{k,j}(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - c_j[f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}.\end{multlined}
%\end{align*}
and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,p_k\), to get the system of equations,
\begin{align}
	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi_k(x)\tr{}T_{ij}\boldsymbol\phi_k (x)\wrt x  
	%
	\\{}+  \int_{x=y_k}^{y_{k+1}}\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi_k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}},\quad k\in\mathbb Z. \end{multlined} \label{eqn:DG system}
\end{align}

\section{Mass, stiffness, and flux matrices}
For \(k\in\mathbb Z\), define local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x.\]
%and global mass and stiffness matrices by
%\begin{align*}
%	M &:= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
%	\\G &:= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
%\end{align*}
We can write (\ref{eqn:DG system}) as 
\begin{align}
	&\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \bs M_k = \sum_{i\in\calS} \boldsymbol a_{k,i}(t)\bs M_kT_{ij} 
	%
	+  c_j\boldsymbol a_{k,j}(t) \bs G_k - c_j[f_j(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}}.\label{eqn:DG matrix}
\end{align}

@@

Let us first consider the boundary at \( X_t=0\). The boundary conditions that describe the evolution of probability/density of a stochastic fluid queue with a boundary at \(0\) are;
\begin{align}\label{eqn:BC1}
\cfrac{\wrt}{\wrt t}\bs q_{-1}(t) &= \bs q_{-1}(t) \bs T_{{-1},{-1}} + \boldsymbol f_-(0,t)\bs C_-\bs P_{-,{-1}},
\\\label{eqn:BC2}
\bs f_-(0^+,t)\bs C_-\bs P_{-,+} + \bs q_{-1}(t)\bs T_{{-1},+}&=\bs f_+(0,t)\bs C_+,
\end{align}
where \(\bs P_{-,+} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{+}}\) and \(\bs P_{-,{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\). Equation~(\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(\bs T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(\boldsymbol f_-(0,t)\bs P_{-,{-1}}\bs C_{-1}\). Substituting the DG approximation for \(\boldsymbol f_-(0,t)\) into (\ref{eqn:BC1}) gives, for \(j\in\calS_{-1}\), 
\begin{equation*}\label{eqn:DGBC1}
\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0^+)\tr{}p_{ij}^{-1} c_i.
\end{equation*}
Equation~(\ref{eqn:BC2}) describes the reflection of process at \(0\), plus the flux of probability mass to density upon a transition from a phase in \(\calS_{-1}\) to a phase in \(\calS_+\). Thus, the flux into the left-hand edge of \(\calD_{0,j}\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_{-1}} q_{{-1},i}(t)T_{ij}\). Therefore, we can now evaluate 
\begin{align*}
	&-c_j[f_j(x,t)\boldsymbol\phi_0(x)]_{x=0}^{x=y_{0}} 
	\\& =  -c_jf_j(y_0^-,t)\boldsymbol\phi_0(y_0)+c_jf_j(0^+,t)\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_if_i(0^+,t)p_{ij}^{-1}\bs \phi_0(0)
	\\& \approx -c_j f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs \phi_0(0)\tr{}\bs \phi_0(0)
	\\& = c_j\boldsymbol a_{0,j}(t)\bs F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs F_j^{0,0},
\end{align*}
for \(j \in \calS_+\).

@@

% APPLICATION TO INTERGRATING OVER TIME
\section{Time-integration schemes}\label{sec: time integration}
In this thesis we numerically integrate ODEs of the form
\begin{align}\label{eqn: DG ODE w BCs 3}
	\cfrac{\wrt}{\wrt t} \boldsymbol a(t)
	% 
	= \boldsymbol a(t) \bs Q
\end{align}
where \(\bs a(t)\) is a vector of coefficients and \(\bs Q\) a matrix, given an initial condition \(\bs a(0)\). To do so, we employ the strong stability preserving Runge-Kutta (SSPRK) scheme of order 4 with 5 stages \citep{sr2002}. SSPRK methods are a family of Runge-Kutta methods which, to quote Section~5.7 of \cite{nodalDGBook}, ``...can be used with advantage for problems with strong shocks and discontinuities, as they guarantee that no additional oscillations are introduced as part of the time-integration process.''

>>>>

In this chapter we introduce the discontinuous Galerkin (DG) method applied to fluid queues to approximate the operator-analytic expressions for fluid-fluid queues in \cite{bo2014}. 
\begin{center}
	\begin{minipage}{0.8\textwidth}
		\textit{Apart from Sections~\ref{sec: other applications} and \ref{sec: limiting and linearity}, this chapter has been taken from Section~4 and 5 of \cite{blnos2022} with only minor changes, such as notation, so that this chapter is consistent with the rest of the thesis. I am a co-author of the paper \cite{blnos2022}. %The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara. I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme. I contributed Sections~4.4 and 5.1. I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6. Appendix~A is also my original work. I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
		}
	\end{minipage}
	\end{center}
% The DG method for B 
    % direct copy paste from the paper
% Problems with DG, non-neg/oscillations
    % solutions: filtering and limiting
% Approximating R: projection, interpolation, cell averages (eluding to QBD-RAP and unif method)
% Approximating Psi (and the limiting distribution, but this isnt really necessary)
    % constructing D
    % solving the matrix-Riccati equation, iteration/algorithm.

@@

% \section{Discontinuous Galerkin Approximation of the Generator of a Fluid Queue}
	% \label{sec:DG}
Discontinuous Galerkin (DG) methods can be used to approximate the solutions to systems of partial differential equations (PDEs). Here we describe how one can apply the DG method to approximate fluid queues and fluid-fluid queues. In finite-element methods we project the partial differential equations onto a set of piecewise polynomial functions. This projection leads to a new system of equations which represent a \textit{weak form} of the original system of PDEs. Next, we approximate the \textit{flux operator} which moves probability from one cell to another. This creates a system of ordinary differential equations which is known as a \emph{semi-discrete system} (discrete in the spatial variable but continuous in time). We can solve the semi-discrete system to approximate transient distributions of the fluid queue or, as we do in Section~\ref{sec:DGSFFM}, we can extract and manipulate an approximation to the generator of the fluid queue to approximate performance measures of fluid-fluid queues. The DG method conserves probability. Here we construct the DG approximation to the matrix of operators \(\mathbb B\) which we use later to construct a DG approximation to \(\mathbb D(s)\) then \(\mathbb\Psi(s)\), and ultimately the limiting distribution of a stochastic fluid-fluid queue.

@@

For each \(k\) we choose \(p_k\) linearly independent functions \(\{\phi^r_k\}_{r=1}^{p_k}\), compactly supported on \(\calD_{k,i}\) (i.e.~\(\phi^r_k(x)=0\) for \(x\notin\calD_{k,i}\)) to form a basis for the space \(W_k\), in which we formulate the approximation. Here, as is standard in DG methods \citep{nodalDGBook}, we take \(\{\phi^r_k\}_{r=1}^{p_k}\) to be the space of polynomials of degree \(p_k-1\). For the sake of illustration, the reader may think of \(\{\phi^r_k\}_{r=1}^{p_k}\) as the Lagrange polynomials, but any polynomial basis can be used. On each cell \(\mathcal D_{k,i}\) we approximate 
\[f_i(x,t)\approx u_{k,i}(x,t)=\sum\limits_{r=1}^{p_k}a_{k,i}^r(t)\phi^r_k(x),\] 
where \(a_{k,i}^r(t)\) are yet-to-be-determined time-dependent coefficients. We refer to \(u_{k,i}\) as the \textit{local} approximation on cell \(k\), while the \textit{global} approximation is given by \(\sum\limits_{k\in\mathbb Z}u_{k,i}\) on the whole domain. %The whole approximation space is \(\bigoplus\limits_{k\in\mathbb Z} W_k\), the direct sum of the spaces \(W_k\), \(k\in\mathbb Z\).

@@

The DG method proceeds by first considering the \textit{weak-formulation} of the PDE, which is constructed from the \textit{strong-form} of the PDE, Equation~(\ref{eqn:pde_density_matrix}). In general, to construct the weak-form we need a set of test functions, say \(W\). Now, take the strong form of the PDE, multiply it by some test function \(\psi(x)\in W\), integrate with respect to \(x\), and apply integration by parts to get 
\begin{align}
\begin{multlined}[t]
	\int_{x\in\mathbb R}\cfrac{\partial}{\partial t} f_j(x,t)\psi(x)\wrt x = \int_{x\in\mathbb R} \sum_{i\in\calS}f_i(x,t)T_{ij}\psi(x)\wrt x 
	%
	+  \int_{x\in\mathbb R} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\psi(x)\wrt x \\{}- [ f_j(x,t)c_j\psi(x)]_{x=-\infty}^{x=\infty}, \end{multlined}\label{eqn:weak form}
\end{align}
for \(j\in\calS\). It is common to choose \(W\) such that \(\psi(-\infty)=\psi(\infty)=0\), in which case the last term on the right is zero. Requiring (\ref{eqn:weak form}) to hold for every \(\psi\in W\) gives the weak-formulation of the PDE. Solutions to (\ref{eqn:weak form}) are known as \textit{weak} solutions and generalise the concept of a solution of the PDE. For example, this may allow discontinuities with respect to \(x\) in the solution -- something which is ill-defined for the strong form.

For the purpose of DG, we take the set of test functions to be \(W = \bigoplus\limits_{k\in\mathbb Z} W^k\) (the direct sum of \(W^k\)), the same as the set of basis functions of our solution space. Proceeding as described above, the weak formulation is 
\begin{align*}
	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS}  f_i(x,t)T_{ij}\phi^r_k(x)\wrt x  
	%
	+  \int_{x=y_k}^{y_{k+1}} f_j(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x \\ {}- [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
\end{align*}
since \(\phi^r_k\) is compactly supported on \(\calD_{k,j}\), for all \(j\in\mathcal S,\,r\in\mathcal N_k\), \(k\in\mathbb Z.\) Now, note that any function \(g(x)\) can be decomposed as \(g(x) = g^{W}(x)+g^\perp(x)\) where \(g^{W}\in W\) and \(g^\perp \in W^\perp\), and \(W^\perp\) is the orthogonal complement of \(W\). Since \(g^\perp\) is orthogonal to \(W\), \(\displaystyle\int_{x}g^\perp(x)\phi^r_k(x)\wrt x=0\) for \(r\in\mathcal N_k,\,k\in\mathbb Z\). Also, note that \(\cfrac{\wrt}{\wrt x}\phi^r_k(x)\in W\). Using this, we can write 
\begin{align*}
	\begin{multlined}[t]\int_{x=y_{k}}^{y_{k+1}}\cfrac{\partial}{\partial t} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)\phi^r_k(x)\wrt x 
	= \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \left(f_i^W(x,t)+f_i^\perp(x,t)\right)T_{ij}\phi^r_k(x)\wrt x  
	%
	\\ {}
	+  \int_{x=y_k}^{y_{k+1}} \left(f_j^W(x,t)+f_j^\perp(x,t)\right)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}, \end{multlined}
\end{align*}
	which is equivalent to
\begin{align}
	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\partial}{\partial t} f_j^W(x,t)\phi^r_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} f_i^W(x,t)T_{ij}\phi^r_k(x)\wrt x  
	%
	\\{} + \int_{x=y_k}^{y_{k+1}} f_j^W(x,t)c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - [f_j(x,t)c_j\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}. \end{multlined} \label{eqn:hash}
\end{align}
Now, \(f_j^W(x,t)\in W\) so, on \(\mathcal D_{k,j}\), it can be expressed as \(u_{k,j}(x,t):=\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\), where the prime, \({}\tr{}\), denotes the transpose, which we now substitute into (\ref{eqn:hash})
%\begin{align*}
%	\begin{multlined}[t]\int_{x\in\calD_k}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi^k(x)\tr{}\phi^r_k(x)\wrt x = \int_{x\in\calD_k}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi^k(x)\tr{}T_{ij}\phi^r_k(x)\wrt x  
%	%
%	\\ {}+  \int_{x\in\calD_k}\boldsymbol a_{k,j}(t) \boldsymbol \phi^k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\phi^r_k(x)\wrt x - c_j[f_j(x,t)\phi^r_k(x)]_{x=y_k}^{x=y_{k+1}}.\end{multlined}
%\end{align*}
and repeat this for all test functions \(\phi^r_k(x)\), \(r=1,...,p_k\), to get the system of equations,
\begin{align}
	\begin{multlined}[t]\int_{x=y_k}^{y_{k+1}}\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x = \int_{x=y_k}^{y_{k+1}}\sum_{i\in\calS} \boldsymbol a_{k,i}(t) \boldsymbol \phi_k(x)\tr{}T_{ij}\boldsymbol\phi_k (x)\wrt x  
	%
	\\{}+  \int_{x=y_k}^{y_{k+1}}\boldsymbol a_{k,j}(t) \boldsymbol \phi_k(x)\tr{}c_j\cfrac{\wrt}{\wrt x}\boldsymbol\phi_k(x)\wrt x - c_j[f_j(x,t) \boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}},\quad k\in\mathbb Z. \end{multlined} \label{eqn:DG system}
\end{align}

\section{Mass, stiffness, and flux matrices}
Recall, for \(k\in\mathbb Z\), we defined local \textit{mass} and \textit{stiffness} matrices \(\bs M_k\) and \(\bs G_k\) by 
\[\bs M_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\boldsymbol \phi_k(x)\wrt x,\quad \bs G_k := \int_{x\in\calD_k}\boldsymbol \phi_k(x)\tr{}\cfrac{\wrt}{\wrt x} \boldsymbol \phi_k(x)\wrt x.\]
%and global mass and stiffness matrices by
%\begin{align*}
%	M &:= \left[\begin{array}{ccc} M_1 & & \\ & \ddots & \\ & & M_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\boldsymbol \phi(x)\wrt x,
%	\\G &:= \left[\begin{array}{ccc} G_1 & & \\ & \ddots & \\ & & G_K \end{array}\right] = \int_{x\inb}\boldsymbol \phi(x)\tr{}\cfrac{\wrt}{\wrt x}\boldsymbol \phi(x)\wrt x,
%\end{align*}
We can write (\ref{eqn:DG system}) as 
\begin{align}
	&\cfrac{\wrt}{\wrt t} \boldsymbol a_{k,j}(t) \bs M_k = \sum_{i\in\calS} \boldsymbol a_{k,i}(t)\bs M_kT_{ij} 
	%
	+  c_j\boldsymbol a_{k,j}(t) \bs G_k - c_j[f_j(x,t)\boldsymbol\phi_k(x)]_{x=y_k}^{x=y_{k+1}}.\label{eqn:DG matrix}
\end{align}

@@

Let us first consider the boundary at \( X_t=0\). The boundary conditions that describe the evolution of probability/density of a stochastic fluid queue with a boundary at \(0\) are;
\begin{align}\label{eqn:BC1}
\cfrac{\wrt}{\wrt t}\bs q_{-1}(t) &= \bs q_{-1}(t) \bs T_{{-1},{-1}} + \boldsymbol f_-(0,t)\bs C_-\bs P_{-,{-1}},
\\\label{eqn:BC2}
\bs f_-(0^+,t)\bs C_-\bs P_{-,+} + \bs q_{-1}(t)\bs T_{{-1},+}&=\bs f_+(0,t)\bs C_+,
\end{align}
where \(\bs P_{-,+} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{+}}\), \(\bs P_{-,{-1}} = \left[p_{ij}^{-1}\right]_{i\in\calS_-,j\in\calS_{-1}}\) and \(p_{ij}^{-1}\) is the probability of instantaneous transition of the phase process from Phase \(i\) to Phase \(j\) upon the process hitting the lower boundary. Equation~(\ref{eqn:BC1}) states that point mass moves between phases according to the sub-generator matrix \(\bs T_{{-1},{-1}}\), and that the flux of probability density into the point masses is \(\boldsymbol f_-(0,t)\bs P_{-,{-1}}\bs C_{-1}\). Substituting the DG approximation for \(\boldsymbol f_-(0,t)\) into (\ref{eqn:BC1}) gives, for \(j\in\calS_{-1}\), 
\begin{equation*}\label{eqn:DGBC1}
\cfrac{\wrt}{\wrt t} q_{{-1},j}(t) = \sum_{i\in\calS_{-1}} q_{{-1},i}(t) T_{i j} - \sum_{i\in\calS_-}\boldsymbol a_{0,i}(t)\boldsymbol \phi_0(0^+)\tr{}p_{ij}^{-1} c_i.
\end{equation*}
Equation~(\ref{eqn:BC2}) describes the reflection of process at \(0\), plus the flux of probability mass to density upon a transition from a phase in \(\calS_{-1}\) to a phase in \(\calS_+\). Thus, the flux into the left-hand edge of \(\calD_{0,j}\) in phase \(j\in\calS_+\) is, \(\sum\limits_{i\in\calS_{-1}} q_{{-1},i}(t)T_{ij}\). Therefore, we can now evaluate 
\begin{align*}
	&-c_j[f_j(x,t)\boldsymbol\phi_0(x)]_{x=0}^{x=y_{0}} 
	\\& =  -c_jf_j(y_0^-,t)\boldsymbol\phi_0(y_0)+c_jf_j(0^+,t)\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_if_i(0^+,t)p_{ij}^{-1}\bs \phi_0(0)
	\\& \approx -c_j f_j^*(y_0,t)\boldsymbol\phi_0(y_0)+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs \phi_0(0)\tr{}\bs \phi_0(0)
	\\& = c_j\boldsymbol a_{0,j}(t)\bs F_j^{0,0}+\sum_{i\in\calS_{-1}}q_{{-1},i}(t)T_{ij}\boldsymbol\phi_0(0) - \sum_{i\in\calS_-}c_ip_{ij}^{-1}\bs a_{0,i}(t)\bs F_j^{0,0},
\end{align*}
for \(j \in \calS_+\).

@@

% APPLICATION TO INTERGRATING OVER TIME
\section{Remarks on slope limiters, linear operators and application to fluid-fluid queues}\label{sec: limiting and linearity}
The process of slope limiting, as described in Section~\ref{sec: slope limiting}, essentially amounts to a non-linear post-processing of the approximate solution after applying the operator \(\bs B\). Thus, the slope-limited scheme arrives at an approximation which is the action of a non-linear operator applied to the initial condition. The fact that the approximate operator is non-linear makes solving the Riccati Equation~(\ref{eqn:RiccatiPsi}) difficult/impossible, whereas there are efficient techniques to solve matrix-Riccati equations \cite{bot08}. Thus, for the application to fluid-fluid queues slope limiting is not a viable option. We could, however, proceed with a vanilla DG scheme (with no limiter) and, when we apply the operator \(\bs \Psi\) to some initial condition and realise an approximation to the distribution of the fluid queue when \(\{Y(t)\}\) first return to its initial level, we could then post-process the solution with a limiter or filter if oscillations are present in the approximation. The application of the limiter/filter is dependent on the initial condition, so the post-processing needs to be done for every initial condition under consideration. If there are many initial distributions to consider, this could result in an increased computational cost. We do not investigate this post-processing approach in this thesis.
